---
title: "benchpro"
author: Joachim Fritscher
date: October 23, 2022
output: 
  html_document:
  toc: true
  toc_float: true
  toc_collapsed: true
  toc_depth: 3
  number_sections: true
  theme: lumen
params:
  input: input
  input_detailed: input_detailed
  meta: meta
  tools: tools
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12)
options(scipen=999)
options(dplyr.summarise.inform = FALSE)
```

```{r libraries, echo=FALSE, include=FALSE, warnings=FALSE}
source("functions.R")
library(gridExtra)
library(cowplot)
library(scales)
```

```{r, echo=FALSE}
# Setting up variables
ranks <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species", "Species Adj", "Strain")
ranks_sp <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
metrics <- c("F1", "Sensitivity", "Precision", "TP", "TN", "FP", "FN", 
             "PearsonCorrelationIntersect", "PearsonCorrelationUnion","SpearmanCorrelationIntersect", "SpearmanCorrelatioUnion", 
             "Bray-Curtis similarity", "BrayCurtisIntersect", "BrayCurtisUnion", "Bray-Curtis Dissimilarity", 
             "L2Intersect", "L2Union", "ShannonDiversity", "ShannonDiversityGold", "ShannonDiversityDiff",
             "ShannonDiversityTP", "ShannonDiversityGoldTP", "ShannonDiversityDiffTP", "RichnessTP", "RichnessGoldTP", "RichnessTPDiff")
```


```{r, echo=FALSE}
run_overview <- TRUE
run_dataset_summary <- TRUE
run_rank_wise <- TRUE
run_adjusted_fpfn <- TRUE
run_overview_fpfn <- TRUE
run_samplewise_fpfn <- TRUE
run_rank_wise_abundance <- TRUE
run_per_tool  <- TRUE
run_overview_div <- FALSE 
```


```{r debug, echo=FALSE}
meta_path <- "/media/fritsche/Extreme_SSD/results/CAMI/HumanToy/Benchmark/meta.xlsx"
stats_path <- "/media/fritsche/Extreme_SSD/results/CAMI/HumanToy/Benchmark/new_all/stats.tsv"
stats_detailed_path <- "/media/fritsche/Extreme_SSD/results/CAMI/HumanToy/Benchmark/new_all/stats_detailed.xlsx"

dataset_test_path <- "/media/fritsche/Extreme_SSD/results/CAMI/HumanToy/Benchmark/new_all/stats_test.tsv"
dataset_detailed_test_path <- "/media/fritsche/Extreme_SSD/results/CAMI/HumanToy/Benchmark/new_all/stats_detailed_test.tsv"

dataset_bracken_path <- "/media/fritsche/Extreme_SSD/results/CAMI/HumanToy/Benchmark/new_all/bracken_test/stats.tsv"
dataset_bracken_detailed_path <- "/media/fritsche/Extreme_SSD/results/CAMI/HumanToy/Benchmark/new_all/bracken_test/stats_detailed.xlsx"

meta_path <- "/media/fritsche/Extreme_SSD/results/CAMI/MouseGut/Benchmark/meta.xlsx"
stats_path <- "/media/fritsche/Extreme_SSD/results/CAMI/MouseGut/Benchmark/stats_output/stats.tsv"
stats_detailed_path <- "/media/fritsche/Extreme_SSD/results/CAMI/MouseGut/Benchmark/stats_output/stats_detailed.xlsx"

# stats_path <- dataset_test_path
# stats_detailed_path <- dataset_detailed_test_path
# 
# meta_path <- "/media/fritsche/Extreme_SSD/results/CAMI/Marine/Benchmark/meta.xlsx"
# stats_path <- "/media/fritsche/Extreme_SSD/results/CAMI/Marine/Benchmark/stats_output/stats.tsv"
# stats_detailed_path <- "/media/fritsche/Extreme_SSD/results/CAMI/Marine/Benchmark/stats_output/stats_detailed.xlsx"
meta_path <- "/media/fritsche/Extreme_SSD/results/CAMI/AllCAMI/Benchmark/meta.xlsx"
stats_path <- "/media/fritsche/Extreme_SSD/results/CAMI/AllCAMI/Benchmark/stats_output/stats.tsv"
stats_detailed_path <- "/media/fritsche/Extreme_SSD/results/CAMI/AllCAMI/Benchmark/stats_output/stats_detailed.xlsx"
tool_color_path <- "/media/fritsche/Extreme_SSD/results/CAMI/tool_color_all.xlsx"

dataset_sub <- NULL

```

# Data analysis {.tabset}
This RMarkdown is for creating reports as part of the benchpro profiler evaluation.

```{r data_gen, echo=TRUE, include=FALSE}
tree_based_analysis <- TRUE
tools <- NULL
tools <- c("bracken", "bracken GTDB", "mOTUs3", "mOTUs3 GTDB", "MetaPhlAn4", "MetaPhlAn4 GTDB", "protal")
if (params$input != "debug" & file.exists(params$input)) {
  stats_path <- params$input
  stats_detailed_path <- params$input_detailed
  meta_path <- params$meta
  tree_based_analysis <- TRUE
  tools <- params$tools
  
  if (!file.exists(stats_path) | !file.exists(stats_detailed_path) | !file.exists(meta_path)) {
    quit()
  }
} else {
  dataset_sub <- "_(Marine|rhizosphere)_(5|6)"
}
test <- read_table(stats_path)
test$metric 

data <- load_data(
  meta_path = meta_path,
  stats_path = stats_path,
  stats_detailed_path = stats_detailed_path,
  tool_color_path = tool_color_path)

if (is.null(tools)) {
  tools <- unique(data$meta$Tool)
} else {
  data$meta <- data$meta %>% filter(Tool %in% tools)
  data$stats <- data$stats %>% filter(tool %in% tools)
  data$stats_detailed <- data$stats_detailed %>% filter(Tool %in% tools)
}

## DEBUG ################
if (!is.null(dataset_sub)) {
  data$stats <- data$stats[grepl(dataset_sub, data$stats$sample),]
  data$stats_detailed <- data$stats_detailed[grepl(dataset_sub, data$stats_detailed$Sample),]
  data$meta <- data$meta[grepl(dataset_sub, data$meta$ID),]
}

data <- generate_data(data)


```
  

```{r overview_div, echo=FALSE, results="asis", eval = run_overview_div}
knitr::knit_child(text=c("# Overview Diversity",
                         "The tools use different taxonomies and thus group species differently. Based on the gold-std profile for each of the tools the following are the number of species in the gold-standard profiles. This is ONLY based on the gold-std profile, not on the tools prediction."))

overview.df <- data$stats_all %>% filter(metric %in% c("TP", "FN") & rank == "Species" & adjusted == FALSE) %>% 
  group_by(tool, dataset) %>%  drop_na() %>%
  summarise(positives = sum(value)) %>% as.data.frame() %>%
  pivot_wider(names_from="dataset", values_from = "positives")
all_column <- if_else(length(unique(ds$dataset)) == 1, unique(ds$dataset)[1], "all")

overview.df$species_richness <- round(overview.df[, all_column] / max(overview.df[, all_column]), digits=4) %>% pull(all)
datatable(overview.df)
```


```{r overview, echo=FALSE, results="asis", eval = run_overview, warnings=FALSE, message=FALSE}
children <- c()
children <- c(children, knitr::knit_child(text=c(paste('# Overview { .tabset .tabset-dropdown }', sep='')), quiet=TRUE))
datasets <- sort(unique(data$stats_all$dataset))
target_metrics <- c("F1", "Sensitivity", "Precision")
for (dset in datasets) {
  if (nrow(data$stats_all %>% 
      filter(metric %in% target_metrics & adjusted & dataset == dset)) == 0) {
    next
  }
  all_stats_adjusted <- data$stats_all %>% 
    filter(metric %in% target_metrics & adjusted & dataset == dset) %>%
    pivot_wider(names_from = "metric", values_from = "value") %>%
    select(sample, tool, Sensitivity, Precision, F1) %>%
    mutate(F1=round(F1, 3), Precision=round(Precision, 3), Sensitivity=round(Sensitivity, 3))
  
  children <- c(children, knitr::knit_child( text=c(paste('## ', dset, sep='')), quiet=TRUE))
  children <-
    c(children, knitr::knit_child(
      text = c(
        "```{r, echo=FALSE, results=\"asis\", fig.width=12, fig.height=12, warning=FALSE}",
        "datatable(all_stats_adjusted)",
        "```"
      ),
      quiet = TRUE
    ))
}
cat(children, sep='\t')
```




<!-- # Overview Predictions { .tabset .tabset-dropdown } -->
```{r , echo=FALSE, results="asis", eval = FALSE, warnings=FALSE, message=FALSE}
options(dplyr.summarise.inform = FALSE)

children <- c()
for (target_rank in rev(RANKS)) {
  contained <- target_rank %in% unique(data$stats$rank)
  if (!contained) next
  children <- c(children, knitr::knit_child( text=c(paste('## ', target_rank, sep='')), quiet=TRUE))
  overview.df <- data$stats %>% filter(metric %in% c("TP", "FN") & rank == target_rank) %>% 
    group_by(tool, dataset) %>%
    summarise(positives= sum(value), TP=sum(value[metric == "TP"])) %>% 
    mutate(Sensitivity=round(TP/positives, 3)) %>% as.data.frame() %>% select(c("tool", "dataset", "Sensitivity"))  %>% 
    pivot_wider(names_from="dataset", values_from = c("Sensitivity"))
  children <-
    c(children, knitr::knit_child(
      text = c(
        "```{r, echo=FALSE, results=\"asis\", fig.width=12, fig.height=12, warning=FALSE}",
        "datatable(overview.df)",
        "```"
      ),
      quiet = TRUE
    ))
}

cat(children, sep='\t')
```


# Data analysis {.tabset}

```{r dataset_summary, eval=run_dataset_summary, echo=FALSE, message=FALSE, warning=FALSE, results="asis", warnings=FALSE}
###############################################################################################
## Rank 


target_metrics <- c("F1", "Sensitivity", "Precision")

n_datasets <- length(unique(data$stats_all$dataset))
height_single <- 0.5+2*n_datasets
children <- c()
children <- c(children, knitr::knit_child( text=c(paste('## Dataset Summary { .tabset .tabset-dropdown }', sep='')), quiet=TRUE))


for (target_dataset in unique(data$stats_all$dataset)) {

  children <- c(children, knitr::knit_child( text=c(paste('### ', target_dataset, " { .tabset .tabset-dropdown }", sep='')), quiet=TRUE))
  
  for (r in rev(sort(unique(data$stats_all$rank)))) {
    rank_df <- data$stats_all %>% 
      filter(rank == r & !is.na(metric) & dataset == target_dataset)
    
    if (nrow(rank_df) == 0) {
      next
    }
  
    # summary_table <- 
    
    summary_table <- rank_df %>% 
      group_by(tool, dataset, metric) %>%
      summarise(mean_value=mean(value), sd_value=sd(value)) %>%
      mutate(label=paste(round(mean_value, 3), "±", round(sd_value, 3), sep='')) %>%
      select(tool, metric, mean_value, label)
    
    
    children <- c(children, knitr::knit_child( text=c(paste('#### ', r, sep='')), quiet=TRUE))
    children <- c(children, knitr::knit_child(
      text = c(
        paste("```{r, echo=FALSE, results=\"asis\", fig.width=", 12, ", fig.height=", height_single, ", warning=FALSE}", sep=''),
        "datatable(summary_table)",
        "```"
      ),
      quiet = TRUE
    ))
  }
  
}
children <-
  c(children, knitr::knit_child(
    text = c(
      "```{r, echo=FALSE, results=\"asis\", fig.width=12, fig.height=12, warning=FALSE}",
      "cat('\n\n\n\n\n\n\n\n\n\n')",
      "```"
    ),
    quiet = TRUE
  ))
cat(children, sep='\n')
```


```{r rank_wise, echo=FALSE, results="asis", eval = run_rank_wise}
###############################################################################################
## Rank 


target_tools <-  c("protal", "MetaPhlAn4 GTDB", "mOTUs3 GTDB")
target_metrics <- c("F1", "Sensitivity", "Precision")

n_datasets <- length(unique(data$stats_all$dataset))
height_single <- 0.5+2*n_datasets
children <- c()
children <- c(children, knitr::knit_child( text=c(paste('## Rank-wise analysis { .tabset .tabset-dropdown }', sep='')), quiet=TRUE))

for (r in rev(sort(unique(data$stats_all$rank)))) {
  rank_df <- data$stats_all %>% filter(rank == r)
  
  if (nrow(rank_df) == 0) {
    next
  }
  
  children <- c(children, knitr::knit_child( text=c(paste('### ', r, sep='')), quiet=TRUE))
  
  summary_plot <- plot_all(rank_df %>% filter(metric %in% target_metrics), r, tool_colors = data$tool_colors)
  
  
  children <- c(children, knitr::knit_child(
    text = c(
      paste("```{r, echo=FALSE, results=\"asis\", fig.width=", 12, ", fig.height=", height_single, ", warning=FALSE}", sep=''),
      "print(summary_plot)",
      "```"
    ),
    quiet = TRUE
  ))
}
children <-
  c(children, knitr::knit_child(
    text = c(
      "```{r, echo=FALSE, results=\"asis\", fig.width=12, fig.height=12, warning=FALSE}",
      "cat('\n\n\n\n\n\n\n\n\n\n')",
      "```"
    ),
    quiet = TRUE
  ))
cat(children, sep='\n')
```



```{r rank_wise_abundance, echo=FALSE, results="asis", eval = run_rank_wise_abundance}
###############################################################################################
## Rank 

target_metrics_intersect <- c("PearsonCorrelationIntersect", "L2Intersect")
target_metrics_union <- c("PearsonCorrelationUnion", "BrayCurtisIntersect", "L2Union")

n_datasets <- length(unique(data$stats_all$dataset))
height_single <- 0.5+2*n_datasets
children <- c()
children <- c(children, knitr::knit_child( text=c(paste('## Rank-wise abundance analysis { .tabset .tabset-dropdown }', sep='')), quiet=TRUE))

for (r in rev(sort(unique(data$stats_all$rank)))) {
  rank_df <- data$stats_all %>% filter(rank == r)
  
  if (nrow(rank_df) == 0) {
    next
  }
  
  children <- c(children, knitr::knit_child( text=c(paste('### ', r, sep='')), quiet=TRUE))
  
  summary_plot_intersect <- rank_df %>% 
    filter(metric %in% target_metrics_intersect) %>% 
    {if(nrow(.) == 0) NULL else .} %>%
    plot_all(r, tool_colors = data$tool_colors)
  summary_plot_union <- rank_df %>% 
    filter(metric %in% target_metrics_union) %>% 
    {if(nrow(.) == 0) NULL else .} %>%
    plot_all(r, tool_colors = data$tool_colors)
  
  children <- c(children, knitr::knit_child(
    text = c(
      paste("```{r, echo=FALSE, results=\"asis\", fig.width=", 12, ", fig.height=", height_single, ", warning=FALSE}", sep=''),
      ifelse(is.null(summary_plot_intersect), "", "print(summary_plot_intersect)"),
      ifelse(is.null(summary_plot_union), "", "print(summary_plot_union)"),
      "```"
    ),
    quiet = TRUE
  ))
}
cat(children, sep='\n')
```


```{r overview_fpfn, echo=FALSE, results="asis", eval=run_overview_fpfn, fig.width=12, fig.height=12, warning=FALSE}

children <- c()
children <- c(children, knitr::knit_child( text=c(paste('## Overview FP/FN { .tabset .tabset-dropdown }', sep='')), quiet=TRUE))

summary_table <- data$stats_detailed_new %>% 
  filter(Rank == "Species") %>% 
  group_by(Sample) %>% count(Type) %>% 
  pivot_wider(names_from=Type, values_from = n, values_fill = 0) %>% 
  as.data.frame()
summary_table$`FP + FN` <- summary_table$FN + summary_table$FP

children <- c(children, knitr::knit_child( text=c(paste('### Species Adj', sep='')), quiet=TRUE))
children <-
  c(children, knitr::knit_child(
    text = c(
      "```{r, echo=FALSE, results=\"asis\", fig.width=12, fig.height=12, warning=FALSE}",
      "datatable(summary_table)",
      "```"
    ),
    quiet = TRUE
  ))

for (r in rev(sort(unique(data$stats_detailed$Rank)))) {
  summary_table <- data$stats_detailed %>% 
    filter(Rank == r) %>% 
    group_by(Sample) %>% count(Type) %>% 
    pivot_wider(names_from=Type, values_from = n, values_fill = 0) %>% 
    as.data.frame()
  if (all(c("FP", "FN") %in% c(colnames(summary_table)))) {
    summary_table$`FP + FN` <- summary_table$FN + summary_table$FP
  }
  
  children <- c(children, knitr::knit_child( text=c(paste('### ', r, sep='')), quiet=TRUE))
  children <-
    c(children, knitr::knit_child(
      text = c(
        "```{r, echo=FALSE, results=\"asis\", fig.width=12, fig.height=12, warning=FALSE}",
        "datatable(summary_table)",
        "```"
      ),
      quiet = TRUE
    ))
}
cat(children, sep='\t')
```


<!-- SAMPLE ADJUSTED FP FN LISTINGS -->

```{r adjusted_fpfn, echo=FALSE, results="asis", eval = run_adjusted_fpfn}
children <- c(knitr::knit_child(text='## Adjusted FP/FN { .tabset .tabset-dropdown }', quiet=TRUE))

tools <- sort(unique(data$stats_detailed_new$Tool))
target_metrics <- c("F1", "Sensitivity", "Precision")

for (tool in tools) {
  display.df <- data$stats_detailed_merged %>% 
    filter(Tool == tool) %>%
    mutate(Abundance=ifelse(Type == "FP" | Type == "FFP", PRED, GOLD)) %>%
    select(Sample, Taxon, Tool, Type, Abundance, Dist, Detectable)
  
  children <- c(children, knitr::knit_child( text=c(paste('### ', tool, sep='')), quiet=TRUE))
  children <- c(children, knitr::knit_child(
    text = c(
      "```{r, echo=FALSE, results=\"asis\", fig.width=12, fig.height=12, warning=FALSE}",
      "datatable(display.df)",
      "```"
    ), quiet=TRUE
  ))
}
cat(children, sep='\n')
```



```{r context_fpfn, echo=FALSE, results="asis", eval=TRUE, fig.width=12, fig.height=12, warning=FALSE}

children <- c(knitr::knit_child( text='## Context FP/FN { .tabset .tabset-dropdown }', quiet=TRUE))

data$fp_neighbors$Sample
for (dataset in sort(unique(data$fp_neighbors$Dataset))) {
  
  fp_neighbors.plot <- data$fp_neighbors %>%
    filter(dataset == Dataset) %>%
    ggplot(aes(x=min_tp, y=GOLD, color=Tool)) +
    facet_wrap(~Tool) +
    xlab("Tree distance to closest TP") +
    ylab("True abundance of closest TP") +
    geom_point(alpha=0.1, size=2)
  
  if (!is.null(data$tool_colors)) {
    fp_neighbors.plot <- fp_neighbors.plot +
      scale_color_manual(values = data$tool_colors)
  }
  
  children <- c(children, knitr::knit_child( text=c(paste('### ', dataset, sep='')), quiet=TRUE))
  children <- c(children, knitr::knit_child(
      text = c(
        "```{r, echo=FALSE, results=\"asis\", fig.width=12, fig.height=6, warning=FALSE}",
        "print(fp_neighbors.plot)",
        "```"
      ),
      quiet = TRUE
    ))
}
cat(children, sep='\n')
```



<!-- Per tool analysis -->

```{r pertool, echo=FALSE, results="asis", eval = run_per_tool}
n_datasets <- length(unique(data$stats_all$dataset))
height_single <- 0.5+1.3*n_datasets
children <- c(knitr::knit_child( text='## Tool-wise analysis { .tabset .tabset-dropdown }', quiet=TRUE))

all_stats_m <- data$stats_all
all_stats_m[all_stats_m$adjusted, "rank"] <- "Species Adj"
tools <- unique(all_stats_m$tool)


for (t in tools) {
  tool_df <- all_stats_m %>% filter(tool == t)
  
  children <- c(children, knitr::knit_child( text=c(paste('### ', t, sep='')), quiet=TRUE))
  
  summary_plot <- plot_tool(tool_df, t)

  children <- c(children, knitr::knit_child(
    text = c(
      paste("```{r, echo=FALSE, results=\"asis\", fig.width=", 12, ", fig.height=", height_single, ", warning=FALSE}", sep=''),
      "print(summary_plot)",
      "```"
    ),
    quiet = TRUE
  ))
}
cat(children, sep='\n')
```

<!--Abundance Threshold analysis -->

```{r abundancethreshold, echo=FALSE, results="asis", eval = TRUE}
children <- c()
children <- c(knitr::knit_child( text='## Abundance Threshold Analysis { .tabset .tabset-dropdown }', quiet=TRUE))

ds.detail.threshold <- data$stats_detailed
ds.detail.threshold$Threshold <- 0
for (threshold in seq(0,0.002, 0.00005)) {
  ds.detail.threshold <- rbind(ds.detail.threshold, apply_abundance_threshold(data$stats_detailed, threshold))
}

ds.detail.adj.threshold <- data$stats_detailed_new
ds.detail.adj.threshold$Threshold <- 0
for (threshold in seq(0,0.002, 0.00005)) {
  ds.detail.adj.threshold <- rbind(ds.detail.adj.threshold, apply_abundance_threshold(data$stats_detailed_new, threshold))
}

ds.stats.threshold <- details_to_binary_stats2(ds.detail.threshold, data$stats)
ds.stats.adj.threshold <- details_to_binary_stats2(ds.detail.adj.threshold, data$stats)

target_dataset <- ifelse(data$has_multiple_datasets, ALL_DATASET, as.character(ds.stats.threshold %>% pull(dataset) %>% head(n=1)))
summary.plot <- plot_abundance_threshold(ds.stats.threshold %>% filter(dataset == target_dataset))
summary.plot.adj <- plot_abundance_threshold(ds.stats.adj.threshold %>% filter(dataset == target_dataset))

children <- c(children, knitr::knit_child( text='### Adjusted Species', quiet=TRUE))
children <- c(children, knitr::knit_child(
    text = c(
      paste("```{r, echo=FALSE, results=\"asis\", warning=FALSE}", sep=''),
      "print(summary.plot.adj)",
      "```"
    ),
    quiet = TRUE
))

children <- c(children, knitr::knit_child( text='### Species', quiet=TRUE))
children <- c(children, knitr::knit_child(
    text = c(
      paste("```{r, echo=FALSE, results=\"asis\", warning=FALSE}", sep=''),
      "print(summary.plot)",
      "```"
    ),
    quiet = TRUE
))

cat(children, sep='\n')
```





```{r samplewise_fpfn, eval=run_samplewise_fpfn, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
children <- c(knitr::knit_child( text='## Sample-wise FP/FN analysis { .tabset }', quiet=TRUE))

for (dataset in unique(data$stats_detailed_merged$Dataset)) {
  children <- c(children, knitr::knit_child( text=c(paste('### ', dataset, ' { .tabset }', sep='')), quiet=TRUE))
  dataset_df <- data$stats_detailed_merged %>% filter(Dataset == dataset)
  for (tool in unique(dataset_df$Tool)) {
    tool_df <- dataset_df %>% filter(Tool == tool)
    dffilter <- dataset
    tlfilter <- tool

    children <- c(children, knitr::knit_child( text=c(paste('#### ', tool, sep='')), quiet=TRUE))
    children <- c(children, knitr::knit_child( text=c(paste('##### Tables: { .tabset .tabset-dropdown }', sep='')), quiet=TRUE))

    # child <- knitr::knit_child( text=c('###### Rank { .tabset }'), quiet=TRUE)
    # children <- c(children, child)
    for (r in rev(unique(data$stats$rank))) {
        children <- c(children, knitr::knit_child( text=c(paste('###### ', r, sep='')), quiet=TRUE))
        summary_df <- data$stats %>%
          filter(dataset == dffilter & tool == tlfilter & rank == r & !is.na(metric)) %>%
          pivot_wider(names_from=metric, values_from=value) %>%
          select (-c(rank)) %>% as.data.frame()
        children <- c(children, knitr::knit_child(
          text = c(
            "```{r, echo=FALSE, results=\"asis\", warning=FALSE}",
            "datatable(summary_df)",
            "```"
          ), quiet=TRUE
        ))
    }

    children <- c(children, knitr::knit_child( text=c(paste('##### Plots: { .tabset .tabset-dropdown }', sep='')), quiet=TRUE))

    for (sample in unique(tool_df$Sample)) {
      children <- c(children, knitr::knit_child( text=c(paste('###### ', sample, sep='')), quiet=TRUE))

      textvar <- paste(dataset, tool, sample)
      subdf <- tool_df %>% filter(Sample == sample)

      child <- knitr::knit_child(input="/usr/users/QIB_fr017/fritsche/Projects/benchpro/knit_sample.Rmd", envir=environment(), quiet = TRUE)
      children <- c(children, child)
    }
    child <- knitr::knit_child( text=c('###### {-}'), quiet=TRUE)
  }
}
cat(children, sep='\n')

```



<!-- ## Tree-based FP/FN analysis { .tabset } -->

```{r treebased, echo=FALSE, results="asis", eval=FALSE}
trees <- na.omit(unique(meta$GoldStdTree))
tree_samples <- meta$ID[!is.na(meta$GoldStdTree)]
tree_tools <- meta$Tool[!is.na(meta$GoldStdTree)]

detail.withtree <- data$stats_detailed %>% filter(Sample %in% tree_samples)

children <- c()
distances.all <- NULL
dist.debug <- list()
for (tree_path in trees) {
  tree <- load_tree(tree_path)

  tools <- as.character(na.omit(unique(meta$Tool[meta$GoldStdTree == tree_path])))
  # print(tools)

  for (tool in tools) {
    children <- c(children, knitr::knit_child( text=c(paste('### ', tool, sep='')), quiet=TRUE))

    detail.df <- detail.withtree %>% filter(Tool == tool & Rank == "Species")
    # browser()
    present_species <- unique(detail.df$Taxon)
    #print(present_species)
    subtree <- keep.tip(tree, present_species[present_species %in% tree$tip.label])
    cophen <- cophenetic(subtree)
    dist.result <- distances(detail.df, cophen)
    dist.debug[[tool]] <- dist.result
    tmp.distances <- dist.result$dist
    tmp.distances$Tool <- tool
    tmp.distances$closest_distance <- min(tmp.distances$distance_to_fn, tmp.distances$distance_to_tp)
    tmp.distances$closest_to <- if_else(tmp.distances$distance_to_fn < tmp.distances$distance_to_tp, "FN", "TP")

    if (is.null(nrow(distances.all))) {
      distances.all <- tmp.distances
    } else {
      distances.all <- rbind(distances.all, tmp.distances)
    }
    tmp.distances$abundance <- if_else(tmp.distances$GOLD == 0, tmp.distances$PRED, tmp.distances$GOLD)
    dist.plot <- tmp.distances %>%
      ggplot(aes(x=distance_to_fn, y=abundance, color=closest_to)) +
      geom_point() +
      ggtitle(paste(tool, "- False Positive and closest tree distance in FN and TP")) +
      scale_x_continuous(trans = scales::pseudo_log_trans(sigma = 0.000001)) +
      scale_y_continuous(trans = scales::pseudo_log_trans(sigma = 0.000001)) +
      theme(aspect.ratio=1)
    children <- c(children, knitr::knit_child( text=c("```{r, echo=FALSE, results=\"asis\"}", "print(dist.plot)", "```"), quiet=TRUE))
    dist.plot2 <- tmp.distances %>%
      ggplot(aes(x=distance_to_fn, y=distance_to_tp, color=closest_to)) +
      geom_point() +
      ggtitle(paste(tool, "- False Positive and closest tree distance in FN and TP")) +
      scale_x_continuous(trans = scales::pseudo_log_trans(sigma = 0.000001)) +
      scale_y_continuous(trans = scales::pseudo_log_trans(sigma = 0.000001)) +
      theme(aspect.ratio=1)
    children <- c(children, knitr::knit_child( text=c("```{r, echo=FALSE, results=\"asis\"}", "print(dist.plot2)", "```"), quiet=TRUE))
  }
}

cat(children, sep='\n')
```




```{r abundance_detection_threshold, echo=FALSE, results="asis", eval=TRUE}


newdet.plot.tp <- data$stats_detailed_new %>%
  filter(Type == "TP" & Rank == "Species" ) %>%
  group_by(Tool) %>%
  mutate(idx = as.integer(with_order(order_by = GOLD, fun = row_number, x = GOLD))) %>%
  filter(idx <= 10) %>%
  mutate(GOLD = GOLD*100) %>%
  ggplot(aes(x=idx, y=GOLD, color=Tool, group=Tool)) +
  geom_line() +
  ggtitle("True positive detection threshold (Species Adj)") +
  ylab("% Relative Abundance") +
  xlab("False Negatives sorted ascending") +
  scale_y_log10() +
  scale_x_continuous(breaks= pretty_breaks())


newdet.plot.fn <- data$stats_detailed_new %>%
  filter(Type == "FN" & Rank == "Species") %>%
  group_by(Tool, Detectable) %>%
  mutate(idx = with_order(order_by = -GOLD, fun = row_number, x = -GOLD)) %>%
  filter(idx <= 10) %>%
  mutate(GOLD = GOLD*100) %>%
  filter(idx < 10) %>%
  ggplot(aes(x=idx, y=GOLD, color=Tool, group=Tool)) +
  geom_line() +
  ggtitle("False negative detection threshold (Species Adj)") +
  ylab("% Relative Abundance") +
  xlab("False Negatives sorted descending") +
  facet_grid(rows=vars(Detectable)) +
  scale_x_continuous(breaks=pretty_breaks())

newdet.plot.fn.unadj <- data$stats_detailed %>%
  filter(Type == "FN" & Rank == "Species") %>%
  group_by(Tool) %>%
  mutate(idx = with_order(order_by = -GOLD, fun = row_number, x = -GOLD)) %>%
  filter(idx <= 10) %>%
  mutate(GOLD = GOLD*100) %>%
  ggplot(aes(x=idx, y=GOLD, color=Tool, group=Tool)) +
  geom_line() +
  ggtitle("False negative detection threshold (Species)") +
  ylab("% Relative Abundance") +
  xlab("False Negatives sorted descending") +
  scale_x_continuous(breaks= pretty_breaks())

newdet.plot.tp.unadj <- data$stats_detailed %>%
  filter(Type == "TP" & Rank == "Species" ) %>%
  group_by(Tool) %>%
  mutate(idx = as.integer(with_order(order_by = GOLD, fun = row_number, x = GOLD))) %>%
  filter(idx <= 10) %>%
  filter(idx <= 10) %>%
  mutate(GOLD = GOLD*100) %>%
  ggplot(aes(x=idx, y=GOLD, color=Tool, group=Tool)) +
  geom_line() +
  ggtitle("True positive detection threshold (Species)") +
  xlab("False Negatives sorted ascending") +
  ylab("% Relative Abundance") +
  scale_x_continuous(breaks= pretty_breaks())


if (!is.null(data$tool_colors)) {
  newdet.plot.tp <- newdet.plot.tp + 
    scale_color_manual(values = data$tool_colors)
  newdet.plot.fn <- newdet.plot.fn + 
    scale_color_manual(values = data$tool_colors)
  newdet.plot.fn.unadj <- newdet.plot.fn.unadj + 
    scale_color_manual(values = data$tool_colors)
  newdet.plot.tp.unadj <- newdet.plot.tp.unadj + 
    scale_color_manual(values = data$tool_colors)
}

children <- c()
children <- c(knitr::knit_child( text='## Abundance detection threshold { .tabset }', quiet=TRUE))
children <- c(children, knitr::knit_child( text='### Adjusted Species', quiet=TRUE))
children <- c(children, knitr::knit_child(
    text = c(
      paste("```{r, echo=FALSE, results=\"asis\", warning=FALSE}", sep=''),
      "print(newdet.plot.tp)",
      "print(newdet.plot.fn)",
      "```"
    ),
    quiet = TRUE
))
children <- c(children, knitr::knit_child( text='### Unadjusted Species', quiet=TRUE))
children <- c(children, knitr::knit_child(
    text = c(
      paste("```{r, echo=FALSE, results=\"asis\", warning=FALSE}", sep=''),
      "print(newdet.plot.tp.unadj)",
      "print(newdet.plot.fn.unadj)",
      "```"
    ),
    quiet = TRUE
))

cat(children, sep='\n')
```



## Alpha diversity detection threshold { .tabset }

```{r, echo=FALSE, results="asis", eval=FALSE}

target_stats <- c("ShannonDiversity", "ShannonDiversityGold", "ShannonDiversityDiff",
                  "ShannonDiversityTP", "ShannonDiversityGoldTP", "ShannonDiversityDiffTP",
                  "RichnessTP", "RichnessGoldTP", "RichnessTPDiff")

#ds %>% filter(rank == r & metric %in% target_stats)

children <- c()
for (r in rev(unique(data$stats$rank))) {
  children <- c(children, knitr::knit_child( text=c(paste('### ', r, sep='')), quiet=TRUE))
  dssub <- data$stats %>% filter(rank == r & metric %in% target_stats)
  dssub$metric <- ordered(dssub$metric, levels=target_stats)
  children <- c(children, knitr::knit_child(input="/usr/users/QIB_fr017/fritsche/Projects/benchpro/knit_alpha_diversity_child.Rmd", envir=environment(), quiet = TRUE))
}
cat(children, sep='\n')
```

# {-}
```{r, echo=FALSE, results="asis", fig.width=12, fig.height=12, warning=FALSE}
plot.new()
```

